Query Optimization for Dynamic Imputation

### 2017-03-01 paper/algorithm.tex
paper: widows & orphans

\section{Algorithm}
In order to correctly plan around missing values, we first extend the relational algebra operators (selection $\sigma$, projection $\pi$, join $\bowtie$, and group-by/aggregate $g$) 
with two new operators (\Cref{sec:operators}). We then define the search space (\Cref{sec:search-space}) of plans to encompass non-trivial operations, such as joins and 
aggregations. Soundness of query execution is provided by our main design invariant, which guarantees traditional relational
algebra operators must never observe a missing value in any attribute that they operate on directly (\Cref{sec:placement}). 
The decision to place imputation operators is driven by our cost model (\Cref{sec:cost-model}), which characterizes a plan based on estimates for the 
quality of the imputations performed and the runtime performance. Our imputation planning algorithm is
agnostic to the type of imputation used (\Cref{sec:imputation}). Finally, we show
that while our algorithm is exponential in the number of joins (\Cref{sec:complexity}), in practice, planning times are not a concern.

@@

The space of query plans is similar to that considered in a canonical
cost-based optimizer~\cite{blasgen1981system}, with the addition of imputation operators appearing before and after traditional operators.
\Cref{fig:query-schematic} shows a plan schematic for a query involving three tables, absent any imputation operators.

@@

\subsection{Complexity}\label{sec:complexity}
Our optimization algorithm builds off the approach taken by 
a canonical cost-base optimizer~\cite{blasgen1981system}.
If imputation operators are ignored, we search the same plan space.
Therefore, our algorithm has a complexity of at least $O(2^J)$, where $J$ is the number of joins.

>>>>

\vspace{0.5in}
\section{Algorithm}
In order to correctly plan around missing values, we first extend the relational algebra operators (selection $\sigma$, projection $\pi$, join $\bowtie$, and group-by/aggregate $g$) 
with two new operators (\Cref{sec:operators}). We then define the search space (\Cref{sec:search-space}) of plans to encompass non-trivial operations, such as joins and 
aggregations. Soundness of query execution is provided by our main design invariant, which guarantees traditional relational
algebra operators must never observe a missing value in any attribute that they operate on directly (\Cref{sec:placement}). 
The decision to place imputation operators is driven by our cost model (\Cref{sec:cost-model}), which characterizes a plan based on estimates for the 
quality of the imputations performed and the runtime performance. Our imputation planning algorithm is
agnostic to the type of imputation used (\Cref{sec:imputation}). Finally, we show
that while our algorithm is exponential in the number of joins (\Cref{sec:complexity}), in practice, planning times are not a concern.

@@

The space of query plans is similar to that considered by System R~\cite{blasgen1981system}, with the addition of imputation operators appearing before and after traditional operators.
\Cref{fig:query-schematic} shows a plan schematic for a query involving three tables, absent any imputation operators.

@@

\subsection{Complexity}\label{sec:complexity}
Our optimization algorithm builds off the approach taken by System R~\cite{blasgen1981system}.
If imputation operators are ignored, we search the same plan space.
Therefore, our algorithm has a complexity of at least $O(2^J)$, where $J$ is the number of joins.
